---
title: "Lab 4"
author: "Trevor Labrecque"
date: "9/29/2020"
output: html_document
---
Load packages
```{r}
library(tidyverse)
library(readxl)
```

Load in the Data
```{r}
f <- file.choose()
states <- read_excel(f)
```

Calculate number of rows and columns in each dataset
```{r}
nrow(dennys)
ncol(dennys)
nrow(laquinta)
ncol(laquinta)
```

Countries For La Quinta
Honduras
Turkey
Chile
Columbia
Canada 
USA 
Mexico

Countries Dennys
USA

Using the filter command to identify which locations have states that dont match abbreviations from the set.
```{r}
dennys %>%
  filter(!(state %in% states$abbreviation))
```
All Dennys locations are in the US


Mutate Dennys dataset to include country column
```{r}
dennys <- dennys %>%
  mutate(country = "United States")
```

```{r}
lq <- laquinta %>%
  mutate(country = case_when(
    state %in% state.abb     ~ "United States",
    state %in% c("ON", "BC") ~ "Canada",
    state == "ANT"           ~ "Colombia",
    state == c("AG", "QR", "NL")            ~ "Mexico",
    
  ))
```

Now filtering to US only
```{r}
laquinta <- lq %>% filter(country == "United States")
```

Counting by state
```{r}
count(dennys, state, sort = TRUE)
count(laquinta, state, sort = TRUE)
```

All seemed to be fairly in order as the larger states had more of each of these.

Bind Dennys and La Quinta
```{r}
dennys <- dennys %>%
  mutate(establishment = "Denny's")
laquinta <- laquinta %>%
  mutate(establishment = "La Quinta")
dn_lq <- bind_rows(dennys, laquinta)
```

Create a plot
```{r}
ggplot(dn_lq, mapping = aes(x = longitude, y = latitude, color = establishment)) +
  geom_point()
```
Filter to NC and Plot
```{r}
dn_lq2 <- dn_lq %>%
  filter(state == "NC")

ggplot(dn_lq2, mapping = aes(x = longitude, y = latitude, color = establishment)) +
  geom_point()

```

Filter to Texas and plot
```{r}
dn_lq3 <- dn_lq %>%
  filter(state == "TX")

ggplot(dn_lq3, mapping = aes(x = longitude, y = latitude, color = establishment)) +
  geom_point()

```