---
title: "Lab 4"
author: "Trevor Labrecque"
date: "9/29/2020"
output: html_document
---
Load packages
```{r}
library(tidyverse)
library(readxl)
```

Load in the Data
```{r}
f <- file.choose()
states <- read_excel(f)
f <- file.choose()
fastFood <- read_excel(f)
```

Filter to Alaska
```{r}
df_ak <- dennys %>%
  filter(state == "AK")
lq_ak <- laquinta %>%
  filter(state == "AK")
```

Join Data Frames for Alaska Dennys and La Quinta data frames
```{r}
dn_lq_ak <- full_join(df_ak, lq_ak, by = "state")
```

Calculates distance between each Dennys and each Laquinta in Alaska
```{r}
dn_lq_ak = dn_lq_ak %>%
  mutate(distance = 
           haversine(longitude.y, latitude.y, longitude.x, latitude.x, round = 3))
```

Finds the Minimum Distance between for each dennys to a La Quinta in Alaska
```{r}
dn_lq_ak_mindist = dn_lq_ak %>%
  group_by(address.x) %>%
  summarise(closest = min(distance))
```



Filter to NC
```{r}
df_nc <- dennys %>%
  filter(state == "NC")
lq_nc <- laquinta %>%
  filter(state == "NC")
```

Join Data Frames for NC Dennys and La Quinta data frames
```{r}
dn_lq_nc <- full_join(df_nc, lq_nc, by = "state")
```

Calculates distance between each Dennys and each Laquinta in NC
```{r}
dn_lq_nc = dn_lq_nc %>%
  mutate(distance = 
           haversine(longitude.x, latitude.x, longitude.y, latitude.y, round = 3))
```

Finds the Minimum Distance between for each dennys to a La Quinta in NC
```{r}
dn_lq_nc_mindist = dn_lq_nc %>%
  group_by(address.x) %>%
  summarise(closest = min(distance))
```



Filter to Texas
```{r}
df_tx <- dennys %>%
  filter(state == "TX")
lq_tx <- laquinta %>%
  filter(state == "TX")
```

Join Data Frames for Texas Dennys and La Quinta data frames
```{r}
dn_lq_tx <- full_join(df_tx, lq_tx, by = "state")
```

Calculates distance between each Dennys and each Laquinta in Texas
```{r}
dn_lq_tx = dn_lq_tx %>%
  mutate(distance = 
           haversine(longitude.x, latitude.x, longitude.y, latitude.y, round = 3))
```

Finds the Minimum Distance between for each dennys to a La Quinta in NC
```{r}
dn_lq_tx_mindist = dn_lq_tx %>%
  group_by(address.x) %>%
  summarise(closest = min(distance))
```

Renames the providence variable to state for easy implementation
```{r}
fastFood %>% rename(state = province)
```

Tests the nearest_restaurant function
```{r}
test = nearest_restaurant('TX')
```

FUNCTIONS********************************************************************

Haversine distance function to calculate the distance between two longituse and latitude coordinates
```{r}
haversine <- function(long1, lat1, long2, lat2, round = 3) {
  # convert to radians
  long1 = long1 * pi / 180
  lat1  = lat1  * pi / 180
  long2 = long2 * pi / 180
  lat2  = lat2  * pi / 180
  
  R = 6371 # Earth mean radius in km
  
  a = sin((lat2 - lat1)/2)^2 + cos(lat1) * cos(lat2) * sin((long2 - long1)/2)^2
  d = R * 2 * asin(sqrt(a))
  
  return( round(d,round) ) # distance in km
}
```

Finds the nearest restaurant to each La Quinta location in a particular state
```{r}
nearest_restaurant = function (this_state) {
  
  fastFood_state <- fastFood %>%
  filter(state == this_state)
  
  lq_state = lq %>%
  filter(state == this_state)
  
  fastFood_lq_state_nearest <- full_join(lq_state, fastFood_state, by = "state") %>%
    mutate(distance=
           haversine(longitude.x, latitude.x, longitude.y, latitude.y, round = 4)) %>%
  group_by(address.x) %>%
  summarise(closest = min(distance))
}

```

Finds minimum distance to a Dennys from each La Quinta location
```{r}
analyze_state_q = function (this_state) {
  
  dn_state <- dennys %>%
  filter(state == this_state)
  
  lq_state = lq %>%
  filter(state == this_state)
  
  dn_lq_state_nearest <- full_join(lq_state, dn_state, by = "state") %>%
    mutate(distance=
           haversine(longitude.x, latitude.x, longitude.y, latitude.y, round = 4)) %>%
  group_by(address.x) %>%
  summarise(closest = min(distance))  
}

```



I find it showed that the list of fast food restaurants was overall closer to a La Quinta than a Dennys. Where Dennys may be the closest of any particular restaurant, having several restaurants in one made it tough for Dennys to compete.


```{r}
all_state_mindist = nearest_restaurant('TX')
all_state_dennys = analyze_state_q('TX')
all_state = full_join(x=all_state_dennys, y=all_state_mindist, by = "address.x") 
all_state <- all_state %>% pivot_longer(2:3)
all_state %>% ggplot() +
  geom_density(aes(x=value, color = name))
```

